Continuous Deductive Reasoning over Streamed RDF Data

### 2015-07-09 content/contribution/continuous-datalog.tex
fixed the use of primes, and added the commutative diagram (as Nick has not yet pushed a copy)

\begin{definition}[Entailment Justification]
\labeldef{continuous datalog: justification}
A justification for some entailment $f$ of some Datalog program
${\Pi = \langle R , F \rangle \vdash f}$ is a minimal set ${J \subseteq F}$
such that ${\Pi' = \langle R , J \rangle \vdash f}$.
\end{definition}

@@

It should be noted that \refdef{continuous datalog: justification} of entailment
justification is equally applicable to time-annotated Datalog programs,
simply replacing entailed and justifying facts with fact instances.
Furthermore, each justification for $f$ in some $F^T$ entails a unique
instance ${f | t}$, such that many instances of $f$ may occur in the set
$E^T$ of fact instances entailed by
${\Pi^T = \langle R, W^{S,x}_i \rangle}$, each with a single
justification.  As such, entailment justification for fact instances may be
defined formally by \refdef{continuous datalog: instance justification}.
\begin{definition}[Instance Entailment Justification]
\labeldef{continuous datalog: instance justification}
	A justification for some entailment ${f|t}$ of some time-annotated Datalog program
	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
	such that ${\Pi^T' = \langle R , J^T \rangle \vdash f|t}$.
	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
\end{definition}

This allows us to formally specify that the time $t$ at which an entailed fact
instance ${f | t}$ becomes true is the last time at which any of its
justifying instances become true:
\[ t = max(\{ t' : f' | t' \in just(\Pi^T, f | t) \}) \]
As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
program ${\langle R, F^T \rangle}$ is a subset of the timestamps for the instances in $F^T$.

@@

As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
Datalog facts, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
constitutes a valid Datalog program that may be evaluated by the truth
function of Datalog $T^{D}$, the result of which is the definition for the
instantaneous truth of $\Pi^C$ at instant $i$ by $T^{CD}$.
\begin{multline*}
T^{CD} \left( \langle R, F, S, r \rangle, i \right) = \\
	T^{D} \left( \langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \} \right)
\end{multline*}

@@

To sum up, a time-annotated Datalog program
${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle}$ may be derived from
a continuous Datalog program ${\Pi^C = \langle R, F, S, r \rangle}$ by
applying a sliding window $W^{S,r}_i$ of range $r$ over the stream $S$ at
instant $i$ and instantiating the facts in $F$ as $F^0$. At the same time,
the entailments of a ${\Pi^C_i \vDash E^0 \cup E^T_i}$ may be derived from
the entailments of the ${\Pi^C \vDash \langle E, P, N \rangle}$ as the union
of the entailments of ${\langle R, F^0 \rangle \vDash E^0}$ and the set
difference of the windows $W^{P,r}_i$ and $W^{N,r}_i$ of range $r$ over
streams $P$ and $N$ at instant $i$.
Furthermore, a conventional Datalog program ${\Pi = \langle R, F \rangle}$
may be derived from a time-annotated Datalog program $\Pi^C_i$ by
${strip}$ping the time annotations from the set of facts
${F^0 \cup W^{S,r}_i}$, as the entailments of ${\Pi \vDash E}$ may be
derived from the entailments of $\Pi^C_i$ by $strip$ping them.
This yields the TODO RMEMBER NAME OF GRAPH in \reffig{CD-TD-D}.
\begin{figure}
	\centering
	% TODO make the diagram
	\caption{TODO}
	\labelfig{CD-TD-D}
\end{figure}
\end{nestedsection}

>>>>

\begin{definition}[Entailment Justification]
\labeldef{continuous datalog: justification}
A justification for some entailment $f$ of some Datalog program
${\Pi = \langle R , F \rangle \vdash f}$ is a minimal set ${J \subseteq F}$
such that ${\Pi^\prime = \langle R , J \rangle \vdash f}$.
\end{definition}

@@

It should be noted that \refdef{continuous datalog: justification} of entailment
justification is equally applicable to time-annotated Datalog programs,
simply replacing entailed and justifying facts with fact instances.
Furthermore, each justification for $f$ in some $F^T$ entails a unique
instance ${f | t}$, such that many instances of $f$ may occur in the set
$E^T$ of fact instances entailed by
${\Pi^T = \langle R, W^{S,x}_i \rangle}$, each with a single
justification.  As such, entailment justification for fact instances may be
defined formally by \refdef{continuous datalog: instance justification}.
\begin{definition}[Instance Entailment Justification]
\labeldef{continuous datalog: instance justification}
	A justification for some entailment ${f|t}$ of some time-annotated Datalog program
	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
	such that ${\Pi^{T\prime} = \langle R , J^T \rangle \vdash f|t}$.
	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
\end{definition}

This allows us to formally specify that the time $t$ at which an entailed fact
instance ${f | t}$ becomes true is the last time at which any of its
justifying instances become true:
\[ t = max(\{ t^\prime : f^\prime | t^\prime \in just(\Pi^T, f | t) \}) \]
As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
program ${\langle R, F^T \rangle}$ is a subset of the timestamps for the instances in $F^T$.

@@

As both $F$ and ${W_i = \{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
Datalog facts, the tuple ${\langle R, F \cup W_i}$
constitutes a valid Datalog program that may be evaluated by the truth
function of Datalog $T^{D}$, the result of which is the definition for the
instantaneous truth of $\Pi^C$ at instant $i$ by $T^{CD}$.
\begin{multline*}
T^{CD} \left( \langle R, F, S, r \rangle, i \right) = \\
	T^{D} \left( \langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \} \right)
\end{multline*}

@@

To sum up, a time-annotated Datalog program
${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle}$ may be derived from
a continuous Datalog program ${\Pi^C = \langle R, F, S, r \rangle}$ by
applying a sliding window $W^{S,r}_i$ of range $r$ over the stream $S$ at
instant $i$ and instantiating the facts in $F$ as $F^0$. At the same time,
the entailments of a ${\Pi^C_i \vDash E^0 \cup E^T_i}$ may be derived from
the entailments of the ${\Pi^C \vDash \langle E, P, N \rangle}$ as the union
of the entailments of ${\langle R, F^0 \rangle \vDash E^0}$ and the set
difference of the windows $W^{P,r}_i$ and $W^{N,r}_i$ of range $r$ over
streams $P$ and $N$ at instant $i$.
Furthermore, a conventional Datalog program ${\Pi = \langle R, F \rangle}$
may be derived from a time-annotated Datalog program $\Pi^C_i$ by
${strip}$ping the time annotations from the set of facts
${F^0 \cup W^{S,r}_i}$, as the entailments of ${\Pi \vDash E}$ may be
derived from the entailments of $\Pi^C_i$ by $strip$ping them.
This yields the commutative diagram in \reffig{CD-TD-D}.
\begin{figure}
	\centering
	$
	\begin{CD}
		{\Pi^C = \langle R, F, S, r \rangle} @>\vDash>> {\langle E, P, N \rangle} \\
		@V{window}VV @VV{window}V \\
		{\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle} @>\vDash>> {E^0 \cup E^T_i} \\
		@V{strip}VV @VV{strip}V \\
		{\Pi = \langle R, F \cup W_i \rangle} @>\vDash>> {E \cup E_i}
	\end{CD}
	$
	\caption{Commutative diagram from Continuous Datalog programs to Datalog entailments.}
	\labelfig{CD-TD-D}
\end{figure}
\end{nestedsection}
